---
title: "Simulations for Manuscript"
author: "Daniel Rud"
date: "2024-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this R markdown, we will include all simulations included in the manuscript *fill in *  


Libraries
```{r}
library(magrittr)
library(ggplot2)
library(grid)
```

Source simulation function files 

```{r}
source("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/functions_for_sims.R")
```

Set the seed that is passed in the simulations 

```{r}
future.seed = 2025
```

Set globally the number of simulation replicates 
```{r}
num_sims = 1000
```

Set globally the learners in the superlearner for the propensity 

```{r}
propensity_SL.library = c(
  "SL.glmnet",
  "SL.bartMachine",
  "SL.rpart"
)

# propensity_SL.library = c(
#   "SL.glmnet",
#   "SL.rpart"
# )

# propensity_SL.library = c("SL.glmnet", "SL.randomForest", "SL.rpart", "SL.bartMachine" )

```

Set globally simulation parameters

```{r}
n_smp = 2000 # sample size
n_snp = 10 # number of SNPs to include in the analysis
exposure_prev = .5 # controls prevalence of exposure -- intercept for propensity 
SNP_MAF = rep(.3, n_snp) # binomial probability for MAF for each SNP 
SNP_adj_effs_0 = c(rep(.4,3), rep(-.3,3), rep(.1,3)) # effects of adjusting SNPs at EM level 0 
SNP_adj_effs_1 = SNP_adj_effs_0 # effects of adjusting SNPs at EM level 1 
SNP_adj_effs_2 = SNP_adj_effs_0 # effects of adjusting SNPs at EM level 2
rho = -1 # level of correlation -- "-1" defines independence, rho in (0,1) = proximity decay correlation 
```

Set globally the options for TMLE 

```{r}
TMLE_args_list = list(
  outcome_method = "glmnet", 
  npv_thresh = (5/sqrt(n_smp))/log(n_smp), 
  near_positivity_method = "trim", 
  nfolds_cv_Q_init = 10, 
  nfolds_cv_glmnet_outcome = 5,
  alpha_outcome = .5, 
  clever_cov_propensity_wt = T
)
```

# Bias Simulations 

Note: RMSE results are incorrect, need to rerun in order to have accurate results.  

## Binary Outcome -- control for MOR

### MOR -- Linear confounding 

```{r}
#### GLOBALS FOR CONFOUNDING ###################################################

# propensity model 
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .02, 
                                  "sex" = -.2,
                                  "cohort2" = .1, 
                                  "cohort3" = -.1, 
                                  "cohort4" = -.2, 
                                  "cohort5" = -.2, 
                                  "ancestry_1" = .1,
                                  "ancestry_2" = -.1)

# outcome model 
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_outcome_effects = c("age" = .015, 
                               "sex" = -.2,
                               "cohort2" = -.2, 
                               "cohort3" = .2, 
                               "cohort4" = -.1, 
                               "cohort5" = -.1, 
                               "ancestry_1" = .15,
                               "ancestry_2" = -.2)
#################################################################################

family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5) -1 # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45) # underlying ACE effects 
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.1,.1,.1,.1,.1,.1,.1,.1)   # logit effects for snps in propensity, 
# if snps_propensity is NA, does not get used


with_progress(bias_lin_conf_MOR <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family,
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          future.seed = future.seed))

output_results(bias_lin_conf_MOR)

bias_lin_conf_MOR$bias_mean_data

# weighted mean 

# apply(bias_lin_conf_MOR$bias_mean_data, MARGIN = 1, 
#       weighted.mean, w = 1 / c(dbinom(0,2,SNP_MAF[1]), dbinom(1,2,SNP_MAF[1]), dbinom(2,2,SNP_MAF[1])))

bias_lin_conf_MOR$MOR_plot
#saveRDS(bias_lin_conf_MOR, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/bias_lin_conf_MOR.RDS")
```

### MOR -- Misspecification Bias 

<!-- ```{r} -->
<!-- confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2 -->

<!-- confounder_propensity_effects = c("age" = .005, -->
<!--                                   "sex" = .6, -->
<!--                                   "cohort2" = .13, -->
<!--                                   "cohort3" = -.2, -->
<!--                                   "cohort4" = .1, -->
<!--                                   "cohort5" = .3, -->
<!--                                   "ancestry_1" = .25, -->
<!--                                   "ancestry_2" = -.2 -->
<!-- ) -->

<!-- # confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 +  -->
<!-- #   I((X1==0) * (sex == 1) * A) + -->
<!-- #   I((X1==1) * (sex == 0) * A) + I((X1==1) * (sex == 1) * A) +  -->
<!-- #   I((X1==2) * (sex == 0) * A) + I((X1==2) * (sex == 1) * A) -->

<!-- confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 +  -->
<!--   I((X1==0) * (sex == 1) * A) + -->
<!--   I((X1==1) * (sex == 1) * A) +  -->
<!--   I((X1==2) * (sex == 1) * A) -->


<!-- confounder_outcome_effects = c("age" = .0035, -->
<!--                                "sex" = .5, -->
<!--                                "cohort2" = -.1, -->
<!--                                "cohort3" = -.1, -->
<!--                                "cohort4" = .1, -->
<!--                                "cohort5" = .1, -->
<!--                                "ancestry_1" = -.1, -->
<!--                                "ancestry_2" = .1, -->
<!--                                "I((X1==0) * (sex == 1) * A)" = 1, -->
<!--                                "I((X1==1) * (sex == 1) * A)" = -.8 , -->
<!--                                "I((X1==2) * (sex == 1) * A)" =  -1.5 -->
<!--                                ) -->


<!-- # # AWESOME SET!!! USE THESE  -->
<!-- # confounder_outcome_effects = c("age" = .0035, -->
<!-- #                                "sex" = -.75, -->
<!-- #                                "cohort2" = -.1, -->
<!-- #                                "cohort3" = -.1, -->
<!-- #                                "cohort4" = .1, -->
<!-- #                                "cohort5" = .1, -->
<!-- #                                "ancestry_1" = -.1, -->
<!-- #                                "ancestry_2" = .1, -->
<!-- #                                "I((X1==0) * (sex == 1) * A)" = 1.5, -->
<!-- #                                "I((X1==1) * (sex == 1) * A)" = -2 , -->
<!-- #                                "I((X1==2) * (sex == 1) * A)" =  -2.5 -->
<!-- #                                ) -->



<!-- # good sim!!!!  -->
<!-- # confounder_outcome_effects = c("age" = .0035, -->
<!-- #                                "sex" = -.75, -->
<!-- #                                "cohort2" = -.1, -->
<!-- #                                "cohort3" = -.1, -->
<!-- #                                "cohort4" = .1, -->
<!-- #                                "cohort5" = .1, -->
<!-- #                                "ancestry_1" = -.1, -->
<!-- #                                "ancestry_2" = .1, -->
<!-- #                                "I((X1==0) * (sex == 1) * A)" = 1.5, -->
<!-- #                                "I((X1==1) * (sex == 1) * A)" = -1.8 , -->
<!-- #                                "I((X1==2) * (sex == 1) * A)" =  -2.9 -->
<!-- #                                ) -->
<!-- #  -->


<!-- family = "binomial" -->
<!-- SNP1_intercept_effs = c(-1.5,-1,-.5) +.5  # baseline effect of SNP 1 at each level 0,1,2 -->
<!-- ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other   -->
<!-- ACE = c(1.0,1.2,1.45)  # underlying ACE effects  -->
<!-- snps_propensity = NA # which SNPs to include in the propensity model,  -->
<!-- # set to NA for no SNPs in propensity -->
<!-- snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity, -->

<!-- # if snps_propensity is NA, does not get used -->

<!-- # tmle oracle specific formulas -->
<!-- propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2  -->

<!-- outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2  -->


<!-- with_progress(misspecification_bias_MOR_new <- tmle_comparison_1snp(num_sims = num_sims,  -->
<!--                                           n_smp = n_smp,  -->
<!--                                           n_snp = n_snp, -->
<!--                                           family = family,  -->
<!--                                           snps_propensity =snps_propensity,  -->
<!--                                           snp_propensity_effects = snp_propensity_effects, -->
<!--                                           exposure_prev = exposure_prev,  -->
<!--                                           SNP_MAF = SNP_MAF, -->
<!--                                           SNP1_intercept_effs = SNP1_intercept_effs,  -->
<!--                                           SNP_adj_effs_0 = SNP_adj_effs_0,  -->
<!--                                           SNP_adj_effs_1 = SNP_adj_effs_1, -->
<!--                                           SNP_adj_effs_2 = SNP_adj_effs_2, -->
<!--                                           ACE_type = ACE_type, -->
<!--                                           ACE = ACE, -->
<!--                                           rho = rho,  -->
<!--                                           TMLE_args_list = TMLE_args_list,  -->
<!--                                           confounder_data_propensity_formula = confounder_data_propensity_formula,  -->
<!--                                           confounder_propensity_effects = confounder_propensity_effects,  -->
<!--                                           confounder_data_outcome_formula = confounder_data_outcome_formula,  -->
<!--                                           confounder_outcome_effects = confounder_outcome_effects,  -->
<!--                                           propensity_SL.library = propensity_SL.library,  -->
<!--                                           future.seed = future.seed)) -->

<!-- output_results(misspecification_bias_MOR_new) -->
<!-- misspecification_bias_MOR_new$bias_mean_data -->
<!-- misspecification_bias_MOR_new$MOR_plot -->

<!-- ``` -->


### MOR -- Misspecification Bias -- Final Calibration

```{r}
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .005,
                                  "sex" = .6,
                                  "cohort2" = .13,
                                  "cohort3" = -.2,
                                  "cohort4" = .1,
                                  "cohort5" = .3,
                                  "ancestry_1" = .25,
                                  "ancestry_2" = -.2
)

confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 + 
  I((X1==0) * (sex == 1) * A) +
  I((X1==1) * (sex == 1) * A) + 
  I((X1==2) * (sex == 1) * A)





confounder_outcome_effects = c("age" = .0035,
                               "sex" = -.75,
                               "cohort2" = -.1,
                               "cohort3" = -.1,
                               "cohort4" = .1,
                               "cohort5" = .1,
                               "ancestry_1" = -.1,
                               "ancestry_2" = .1,
                               "I((X1==0) * (sex == 1) * A)" = 1.5,
                               "I((X1==1) * (sex == 1) * A)" = -1.0 ,
                               "I((X1==2) * (sex == 1) * A)" =  -1.5
                               )



family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5) +.5  # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45)  # underlying ACE effects 
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,

# if snps_propensity is NA, does not get used

with_progress(misspecification_bias_MOR <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          future.seed = future.seed))

output_results(misspecification_bias_MOR)
misspecification_bias_MOR$bias_mean_data
misspecification_bias_MOR$MOR_plot
#saveRDS(misspecification_bias_MOR, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/misspecification_bias_MOR.RDS")

```

```{r}
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .005,
                                  "sex" = .6,
                                  "cohort2" = .13,
                                  "cohort3" = -.2,
                                  "cohort4" = .1,
                                  "cohort5" = .3,
                                  "ancestry_1" = .25,
                                  "ancestry_2" = -.2
)

confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 + 
  I((X1==0) * (sex == 1) * A) +
  I((X1==1) * (sex == 1) * A) + 
  I((X1==2) * (sex == 1) * A)





confounder_outcome_effects = c("age" = .0035,
                               "sex" = -.75,
                               "cohort2" = -.1,
                               "cohort3" = -.1,
                               "cohort4" = .1,
                               "cohort5" = .1,
                               "ancestry_1" = -.1,
                               "ancestry_2" = .1,
                               "I((X1==0) * (sex == 1) * A)" = -1.5,
                               "I((X1==1) * (sex == 1) * A)" = 1.0 ,
                               "I((X1==2) * (sex == 1) * A)" =  1.5
                               )



family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5) +.5  # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45)  # underlying ACE effects 
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,

# if snps_propensity is NA, does not get used

with_progress(misspecification_bias_MOR_new <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          future.seed = future.seed))

output_results(misspecification_bias_MOR_new)
misspecification_bias_MOR_new$bias_mean_data
misspecification_bias_MOR_new$MOR_plot

```



### MOR -- Omitted Variable Bias 

```{r}
# propensity model
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .001, 
                                  "sex" = .2,
                                  "cohort2" = -.2, 
                                  "cohort3" = -.1, 
                                  "cohort4" = .2, 
                                  "cohort5" = .1, 
                                  "ancestry_1" = -.2,
                                  "ancestry_2" = -.1)



# outcome model
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 

confounder_outcome_effects = c("age" = .002,
                               "sex" = .3,
                               "cohort2" = .3,
                               "cohort3" = .1,
                               "cohort4" = -.2,
                               "cohort5" = .1,
                               "ancestry_1" = .25,
                               "ancestry_2" = -.1
)


family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5)   # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45)  # underlying ACE effects 
snps_propensity = 1:8 # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.5,.5,.5,.5,-.5,-.5,-.5,-.5)   # logit effects for snps in propensity,
# if snps_propensity is NA, does not get used

with_progress(omitted_var_bias_MOR <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library,
                                          future.seed = future.seed))


output_results(omitted_var_bias_MOR)
omitted_var_bias_MOR$bias_mean_data

apply(omitted_var_bias_MOR$bias_mean_data, MARGIN = 1,
      weighted.mean, w = 1 / c(dbinom(0,2,SNP_MAF[1]), dbinom(1,2,SNP_MAF[1]), dbinom(2,2,SNP_MAF[1])))

#saveRDS(omitted_var_bias_MOR, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/omitted_var_bias_MOR.RDS")


```
# Create Trio Plot 

```{r}
lin_plot = bias_lin_conf_MOR$MOR_plot; miss_plot = misspecification_bias_MOR$MOR_plot;  omit_plot = omitted_var_bias_MOR$MOR_plot
miss_plot = miss_plot + ylab(""); omit_plot = omit_plot + ylab("")
lin_plot = lin_plot + xlab(""); omit_plot = omit_plot + xlab("")
lin_plot = lin_plot + ylab("MOR Bias") 
miss_plot = miss_plot + xlab(TeX(r"($S_1 = 0, 1, 2)"))

# add theme for pretty print

lin_plot = lin_plot + theme_classic() + theme(legend.position = "bottom", 
            axis.title = element_text(size = 16),   # Increase axis title size
            axis.text = element_text(size = 12),    # Increase axis labels size
            legend.title = element_text(size = 16), # Increase legend title size
            legend.text = element_text(size = 14)   # Increase legend labels size
            )
miss_plot = miss_plot + theme_classic() + theme(legend.position = "bottom", 
            axis.title = element_text(size = 16),   # Increase axis title size
            axis.text = element_text(size = 12),    # Increase axis labels size
            legend.title = element_text(size = 16), # Increase legend title size
            legend.text = element_text(size = 14)   # Increase legend labels size
            )

omit_plot = omit_plot + theme_classic() + theme(legend.position = "bottom", 
            axis.title = element_text(size = 16),   # Increase axis title size
            axis.text = element_text(size = 12),    # Increase axis labels size
            legend.title = element_text(size = 16), # Increase legend title size
            legend.text = element_text(size = 14)   # Increase legend labels size
            )

pdf(file = "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/Figures/bias_plots.pdf", width = 8, height = 6)
ggarrange(lin_plot, miss_plot, omit_plot,
          nrow = 1,  common.legend = TRUE, legend = "bottom", 
          align = "h", labels = c("A", "B", "C"), 
           label.y = .95)
dev.off()


```



# Power Simulations 

## Binary Outcome -- control for MOR

### MOR -- Linear confounding -- Multiplicative MOR 

```{r}
# propensity model 
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .02, 
                                  "sex" = -.2,
                                  "cohort2" = .1, 
                                  "cohort3" = -.1, 
                                  "cohort4" = -.2, 
                                  "cohort5" = -.2, 
                                  "ancestry_1" = .1,
                                  "ancestry_2" = -.1)

# outcome model 
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_outcome_effects = c("age" = .015, 
                               "sex" = -.2,
                               "cohort2" = -.2, 
                               "cohort3" = .2, 
                               "cohort4" = -.1, 
                               "cohort5" = -.1, 
                               "ancestry_1" = .15,
                               "ancestry_2" = -.2)
#################################################################################
family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5)   # baseline effect of SNP 1 at each level 0,1,2
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,
# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2


ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  

# multiplicative MOR 
u1 = c(1, 1.1,1.2,1.3,1.35, 1.4, 1.45)
ACEs = matrix(0, nrow = length(u1), ncol = 3)
for(i in 1:length(u1))
{
  ACEs[i,] = c(1, u1[i], u1[i]^2)
}

power_lin_conf_mult_MOR = vector(mode = "list", length = length(u1))

for(i in 1: length(u1))
{
  with_progress(power_lin_conf_mult_MOR[[i]] <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACEs[i,],
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          propensity_formula = propensity_formula, 
                                          outcome_formula = outcome_formula,
                                          future.seed = future.seed))
  
  cat(paste0("\nSim",i))
}


MOR_levels_1 = ACEs


process_power_plots_MOR(power_lin_conf_mult_MOR,MOR_levels_1, 
                        log_ACE = F)


#saveRDS(power_lin_conf_mult_MOR, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/power_lin_conf_mult_MOR.RDS")


```



```{r}
# E <- C 
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .005,
                                  "sex" = .6,
                                  "cohort2" = .13,
                                  "cohort3" = -.2,
                                  "cohort4" = .1,
                                  "cohort5" = .3,
                                  "ancestry_1" = .25,
                                  "ancestry_2" = -.2
)

# confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 + 
#   I((X1==0) * (sex == 1) * A) +
#   I((X1==1) * (sex == 0) * A) + I((X1==1) * (sex == 1) * A) + 
#   I((X1==2) * (sex == 0) * A) + I((X1==2) * (sex == 1) * A)

# Y <- C
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 + 
  I((X1==0) * (sex == 1) * A) +
  I((X1==1) * (sex == 1) * A) + 
  I((X1==2) * (sex == 1) * A)

confounder_outcome_effects = c("age" = .0035,
                               "sex" = -.75,
                               "cohort2" = -.1,
                               "cohort3" = -.1,
                               "cohort4" = .1,
                               "cohort5" = .1,
                               "ancestry_1" = -.1,
                               "ancestry_2" = .1,
                               "I((X1==0) * (sex == 1) * A)" = 1.5,
                               "I((X1==1) * (sex == 1) * A)" = -1.5 ,
                               "I((X1==2) * (sex == 1) * A)" =  -2
                               )


# AWESOME SET!!! USE THESE 
# confounder_outcome_effects = c("age" = .0035,
#                                "sex" = -.75,
#                                "cohort2" = -.1,
#                                "cohort3" = -.1,
#                                "cohort4" = .1,
#                                "cohort5" = .1,
#                                "ancestry_1" = -.1,
#                                "ancestry_2" = .1,
#                                "I((X1==0) * (sex == 1) * A)" = 2.5,
#                                "I((X1==1) * (sex == 1) * A)" = -3 ,
#                                "I((X1==2) * (sex == 1) * A)" =  -4.5
#                                )

# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2 

#################################################################################
family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5) + .5   # baseline effect of SNP 1 at each level 0,1,2
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,


ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  

# multiplicative MOR 
u1 = c(1, 1.1,1.2,1.3,1.35, 1.4, 1.45)
ACEs = matrix(0, nrow = length(u1), ncol = 3)
for(i in 1:length(u1))
{
  ACEs[i,] = c(1, u1[i], u1[i]^2)
}

power_lin_conf_mult_MOR_conf = vector(mode = "list", length = length(u1))

for(i in 1: length(u1))
{
  with_progress(power_lin_conf_mult_MOR_conf[[i]] <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACEs[i,],
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          future.seed = future.seed))
  
  cat(paste0("\nSim",i))
}

# process_power_plots(power_lin_conf_mult_MOR_conf, ACE = "MOR")
MOR_levels_2 = ACEs


process_power_plots_MOR(power_lin_conf_mult_MOR_conf,MOR_levels_2, 
                        log_ACE = F)

#saveRDS(power_lin_conf_mult_MOR_conf, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/power_lin_conf_mult_MOR_conf.RDS")

```

Omitted Variable Case 
```{r}
# propensity model
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .001, 
                                  "sex" = .2,
                                  "cohort2" = -.2, 
                                  "cohort3" = -.1, 
                                  "cohort4" = .2, 
                                  "cohort5" = .1, 
                                  "ancestry_1" = -.2,
                                  "ancestry_2" = -.1
                                    )



# outcome model
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2  + I()

confounder_outcome_effects = c("age" = .002,
                               "sex" = .3,
                               "cohort2" = .3,
                               "cohort3" = .1,
                               "cohort4" = -.2,
                               "cohort5" = .1,
                               "ancestry_1" = .25,
                               "ancestry_2" = -.1
)


family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5)   # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45)  # underlying ACE effects 
snps_propensity = 1:8 # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,

# if snps_propensity is NA, does not get used

# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2 


#################################################################################


ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  

# multiplicative MOR 
u1 = c(1, 1.1,1.2,1.3,1.35, 1.4, 1.45)
ACEs = matrix(0, nrow = length(u1), ncol = 3)
for(i in 1:length(u1))
{
  ACEs[i,] = c(1, u1[i], u1[i]^2)
}

power_ommitted_var = vector(mode = "list", length = length(u1))

for(i in 1: length(u1))
{
  with_progress(power_ommitted_var[[i]] <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACEs[i,],
                                          rho = rho, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          future.seed = future.seed))
  
  cat(paste0("\nSim",i))
}

# process_power_plots(power_lin_conf_mult_MOR_conf, ACE = "MOR")
MOR_levels_3 = ACEs


process_power_plots_MOR(power_ommitted_var,MOR_levels_3[1:7,], 
                        log_ACE = F)

#saveRDS(power_ommitted_var, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/power_ommitted_var.RDS")

```

```{r}
#save.image("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/Simulation_results_10-2-24")
load("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/Simulation_results_10-2-24")
```


Save plots from power simulations 
```{r}

# linear confounding 
pdf(file = "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/Figures/power_plot_no_resid_conf.pdf", width = 8, height = 6)
process_power_plots_MOR(power_lin_conf_mult_MOR,MOR_levels_1, 
                        log_ACE = F)
dev.off()

# nonlinear confounding 

pdf(file = "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/Figures/power_plot_conf.pdf", width = 8, height = 6)
process_power_plots_MOR(power_lin_conf_mult_MOR_conf,MOR_levels_1, 
                        log_ACE = F)
dev.off()

```

Make latex tables 

```{r}
generate_latex_table(bias_lin_conf_MOR)

generate_latex_table(misspecification_bias_MOR)

generate_latex_table(omitted_var_bias_MOR)

```


