---
title: "Simulations for Manuscript"
author: "Daniel Rud"
date: "2024-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this R markdown, we will include all simulations included in the manuscript *fill in *  


Libraries
```{r}
library(magrittr)
```

Source simulation function files 

```{r}
source("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/functions_for_sims.R")
```

Set the seed that is passed in the simulations 

```{r}
future.seed = 2024
```

Set globally the number of simulation replicates 
```{r}
num_sims = 1000
```

Set globally the learners in the superlearner for the propensity 

```{r}
propensity_SL.library = c(
  "SL.glmnet",
  "SL.bartMachine",
  "SL.rpart"
)

# propensity_SL.library = c(
#   "SL.glmnet",
#   "SL.rpart"
# )

# propensity_SL.library = c("SL.glmnet", "SL.randomForest", "SL.rpart", "SL.bartMachine" )

```

Set globally simulation parameters

```{r}
n_smp = 2000 # sample size
n_snp = 10 # number of SNPs to include in the analysis
exposure_prev = .5 # controls prevalence of exposure -- intercept for propensity 
SNP_MAF = rep(.3, n_snp) # binomial probability for MAF for each SNP 
SNP_adj_effs_0 = c(rep(.4,3), rep(-.3,3), rep(.1,3)) # effects of adjusting SNPs at EM level 0 
SNP_adj_effs_1 = SNP_adj_effs_0 # effects of adjusting SNPs at EM level 1 
SNP_adj_effs_2 = SNP_adj_effs_0 # effects of adjusting SNPs at EM level 2
rho = -1 # level of correlation -- "-1" defines independence, rho in (0,1) = proximity decay correlation 
W_exposure_null = F # set to `F` to adjust for other SNPs in propensity model for TMLE
W_outcome_null = F # set to `F` to adjust for other SNPs in outcome model for TMLE
```

Set globally the options for TMLE 

```{r}
TMLE_args_list = list(
  outcome_method = "glmnet_int", 
  npv_thresh = (5/sqrt(n_smp))/log(n_smp), 
  near_positivity_method = "trim", 
  nfolds_cv_Q_init = 10, 
  nfolds_cv_glmnet_outcome = 3,
  alpha_outcome = .5, 
  clever_cov_propensity_wt = T,
  outcome_SL.library = NULL, 
  outcome_SL.cvControl = NULL
)
```

# Bias Simulations 

## Binary Outcome -- control for MOR

### MOR -- Linear confounding 

```{r}
#### GLOBALS FOR CONFOUNDING ###################################################

# propensity model 
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .02, 
                                  "sex" = -.2,
                                  "cohort2" = .1, 
                                  "cohort3" = -.1, 
                                  "cohort4" = -.2, 
                                  "cohort5" = -.2, 
                                  "ancestry_1" = .1,
                                  "ancestry_2" = -.1)

# outcome model 
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_outcome_effects = c("age" = .015, 
                               "sex" = -.2,
                               "cohort2" = -.2, 
                               "cohort3" = .2, 
                               "cohort4" = -.1, 
                               "cohort5" = -.1, 
                               "ancestry_1" = .15,
                               "ancestry_2" = -.2)
#################################################################################

family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5) -1 # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45) # underlying ACE effects 
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.1,.1,.1,.1,.1,.1,.1,.1)   # logit effects for snps in propensity, 
# if snps_propensity is NA, does not get used

# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2


with_progress(bias_lin_conf_MOR <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family,
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          W_exposure_null = W_exposure_null, 
                                          W_outcome_null = W_outcome_null, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          propensity_formula = propensity_formula, 
                                          outcome_formula = outcome_formula, 
                                          future.seed = future.seed))

output_results(bias_lin_conf_MOR)

bias_lin_conf_MOR$bias_mean_data

# weighted mean 

apply(bias_lin_conf_MOR$bias_mean_data, MARGIN = 1, 
      weighted.mean, w = 1 / c(dbinom(0,2,SNP_MAF[1]), dbinom(1,2,SNP_MAF[1]), dbinom(2,2,SNP_MAF[1])))


```

### MOR -- Misspecification Bias 

```{r}
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .005,
                                  "sex" = .2,
                                  "cohort2" = .13,
                                  "cohort3" = -.2,
                                  "cohort4" = .1,
                                  "cohort5" = .3,
                                  "ancestry_1" = .25,
                                  "ancestry_2" = -.2
)

confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 + I(sex*(X1 == 1)) + I(sex*(X1 == 2))

confounder_outcome_effects = c("age" = .0035,
                               "sex" = .43,
                               "cohort2" = -.1,
                               "cohort3" = -.1,
                               "cohort4" = .1,
                               "cohort5" = .1,
                               "ancestry_1" = -.1,
                               "ancestry_2" = .1,
                               "I(sex*(X1 == 1))" = -1,
                               "I(sex*(X1 == 2))" = 2.33
                               )

# confounder_outcome_effects = c("age" = .0035,
#                                "sex" = .14,
#                                "cohort2" = -.1,
#                                "cohort3" = -.1,
#                                "cohort4" = .1,
#                                "cohort5" = .1,
#                                "ancestry_1" = -.1,
#                                "ancestry_2" = .1,
#                                "I(sex*(X1 == 1))" = .16,
#                                "I(sex*(X1 == 2))" = -1.5
#                                )



family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5) -1   # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45)  # underlying ACE effects 
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,

# if snps_propensity is NA, does not get used

# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2 

outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2 


with_progress(misspecification_bias_MOR <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          W_exposure_null = W_exposure_null, 
                                          W_outcome_null = W_outcome_null, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          propensity_formula = propensity_formula, 
                                          outcome_formula = outcome_formula, 
                                          future.seed = future.seed))

output_results(misspecification_bias_MOR)
misspecification_bias_MOR$bias_mean_data

# weighted mean

apply(misspecification_bias_MOR$bias_mean_data, MARGIN = 1,
      weighted.mean, w = 1 / c(dbinom(0,2,SNP_MAF[1]), dbinom(1,2,SNP_MAF[1]), dbinom(2,2,SNP_MAF[1])))

# Important note: the ACE bias may be nondecreasing as a function of the SNP level -- this is due to the prevalence of Y within each level X1 = 0,1,2 being different!
# > prop.table(table(data$Y, data$X1), margin= 2)
#
#             0         1         2
#   0 0.6653061 0.5845238 0.4055556
#   1 0.3346939 0.4154762 0.5944444

```



### MOR -- Omitted Variable Bias 

```{r}
# propensity model
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .001, 
                                  "sex" = .2,
                                  "cohort2" = -.2, 
                                  "cohort3" = -.1, 
                                  "cohort4" = .2, 
                                  "cohort5" = .1, 
                                  "ancestry_1" = -.2,
                                  "ancestry_2" = -.1
                                    )



# outcome model
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 

confounder_outcome_effects = c("age" = .002,
                               "sex" = .3,
                               "cohort2" = .3,
                               "cohort3" = .1,
                               "cohort4" = -.2,
                               "cohort5" = .1,
                               "ancestry_1" = .25,
                               "ancestry_2" = -.1
)


family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5)   # baseline effect of SNP 1 at each level 0,1,2
ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  
ACE = c(1.0,1.2,1.45)  # underlying ACE effects 
snps_propensity = 1:8 # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,

# if snps_propensity is NA, does not get used

# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2 

with_progress(omitted_var_bias_MOR <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACE,
                                          rho = rho, 
                                          W_exposure_null = W_exposure_null, 
                                          W_outcome_null = W_outcome_null, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          propensity_formula = propensity_formula, 
                                          outcome_formula = outcome_formula, 
                                          future.seed = future.seed))


output_results(omitted_var_bias_MOR)
omitted_var_bias_MOR$bias_mean_data

apply(omitted_var_bias_MOR$bias_mean_data, MARGIN = 1,
      weighted.mean, w = 1 / c(dbinom(0,2,SNP_MAF[1]), dbinom(1,2,SNP_MAF[1]), dbinom(2,2,SNP_MAF[1])))

```


# Power Simulations 

## Binary Outcome -- control for MOR

### MOR -- Linear confounding -- Multiplicative MOR 

```{r}
# propensity model 
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .02, 
                                  "sex" = -.2,
                                  "cohort2" = .1, 
                                  "cohort3" = -.1, 
                                  "cohort4" = -.2, 
                                  "cohort5" = -.2, 
                                  "ancestry_1" = .1,
                                  "ancestry_2" = -.1)

# outcome model 
confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_outcome_effects = c("age" = .015, 
                               "sex" = -.2,
                               "cohort2" = -.2, 
                               "cohort3" = .2, 
                               "cohort4" = -.1, 
                               "cohort5" = -.1, 
                               "ancestry_1" = .15,
                               "ancestry_2" = -.2)
#################################################################################
family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5)   # baseline effect of SNP 1 at each level 0,1,2
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,
# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2


ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  

# multiplicative MOR 
u1 = c(1, 1.1,1.2,1.3,1.35, 1.4, 1.45)
ACEs = matrix(0, nrow = length(u1), ncol = 3)
for(i in 1:length(u1))
{
  ACEs[i,] = c(1, u1[i], u1[i]^2)
}

power_lin_conf_mult_MOR = vector(mode = "list", length = length(u1))

for(i in 1: length(u1))
{
  with_progress(power_lin_conf_mult_MOR[[i]] <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACEs[i,],
                                          rho = rho, 
                                          W_exposure_null = W_exposure_null, 
                                          W_outcome_null = W_outcome_null, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          propensity_formula = propensity_formula, 
                                          outcome_formula = outcome_formula,
                                          future.seed = future.seed))
  
  cat(paste0("\nSim",i))
}


MOR_levels_1 = ACEs


process_power_plots_MOR(power_lin_conf_mult_MOR,MOR_levels_1, 
                        log_ACE = F)


```



```{r}
# propensity model
confounder_data_propensity_formula = ~ age + sex + cohort + ancestry_1 +  ancestry_2

confounder_propensity_effects = c("age" = .005,
                                  "sex" = .2,
                                  "cohort2" = .13,
                                  "cohort3" = -.2,
                                  "cohort4" = .1,
                                  "cohort5" = .3,
                                  "ancestry_1" = .25,
                                  "ancestry_2" = -.2
)

confounder_data_outcome_formula = ~ age + sex + cohort + ancestry_1 + ancestry_2 + I(sex*(X1 == 1)) + I(sex*(X1 == 2))

confounder_outcome_effects = c("age" = .0035,
                               "sex" = .43,
                               "cohort2" = -.1,
                               "cohort3" = -.1,
                               "cohort4" = .1,
                               "cohort5" = .1,
                               "ancestry_1" = -.1,
                               "ancestry_2" = .1,
                               "I(sex*(X1 == 1))" = -1,
                               "I(sex*(X1 == 2))" = 2.33
                               )


# tmle oracle specific formulas
propensity_formula = A ~ age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2
outcome_formula = Y ~ A + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10 + age + sex + cohort2 + cohort3 + cohort4 + cohort5 + ancestry_1 +  ancestry_2 

#################################################################################
family = "binomial"
SNP1_intercept_effs = c(-1.5,-1,-.5)   # baseline effect of SNP 1 at each level 0,1,2
snps_propensity = NA # which SNPs to include in the propensity model, 
# set to NA for no SNPs in propensity
snp_propensity_effects = c(.2,.2,.2,.2,-.2,-.2,-.2,-.2)   # logit effects for snps in propensity,


ACE_type = "MOR" # ACE type we choose to generate data -- we can control one or other  

# multiplicative MOR 
u1 = c(1, 1.1,1.2,1.3,1.35, 1.4, 1.45)
ACEs = matrix(0, nrow = length(u1), ncol = 3)
for(i in 1:length(u1))
{
  ACEs[i,] = c(1, u1[i], u1[i]^2)
}

power_lin_conf_mult_MOR_conf = vector(mode = "list", length = length(u1))

for(i in 1: length(u1))
{
  with_progress(power_lin_conf_mult_MOR_conf[[i]] <- tmle_comparison_1snp(num_sims = num_sims, 
                                          n_smp = n_smp, 
                                          n_snp = n_snp,
                                          family = family, 
                                          snps_propensity =snps_propensity, 
                                          snp_propensity_effects = snp_propensity_effects,
                                          exposure_prev = exposure_prev, 
                                          SNP_MAF = SNP_MAF,
                                          SNP1_intercept_effs = SNP1_intercept_effs, 
                                          SNP_adj_effs_0 = SNP_adj_effs_0, 
                                          SNP_adj_effs_1 = SNP_adj_effs_1,
                                          SNP_adj_effs_2 = SNP_adj_effs_2,
                                          ACE_type = ACE_type,
                                          ACE = ACEs[i,],
                                          rho = rho, 
                                          W_exposure_null = W_exposure_null, 
                                          W_outcome_null = W_outcome_null, 
                                          TMLE_args_list = TMLE_args_list, 
                                          confounder_data_propensity_formula = confounder_data_propensity_formula, 
                                          confounder_propensity_effects = confounder_propensity_effects, 
                                          confounder_data_outcome_formula = confounder_data_outcome_formula, 
                                          confounder_outcome_effects = confounder_outcome_effects, 
                                          propensity_SL.library = propensity_SL.library, 
                                          future.seed = future.seed))
  
  cat(paste0("\nSim",i))
}

# process_power_plots(power_lin_conf_mult_MOR_conf, ACE = "MOR")
MOR_levels_2 = ACEs


process_power_plots_MOR(power_lin_conf_mult_MOR_conf,MOR_levels_2, 
                        log_ACE = F)

```















```{r}
save.image("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/Simulation_results_10-2-24")
load("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/tlgxe_simulations/Simulation_results_10-2-24")
```