---
title: "UKB Data Analysis"
author: "Daniel Rud"
date: "2024-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries 

```{r}
if (!requireNamespace("tlGxE", quietly = TRUE)) devtools::install_github("Daniel-Rud/tlGxE")
library(tlGxE)
library(magrittr)
```

# Process Data 
Pull files 

```{r}
# original dataset with epi data and 3k SNPs
snp_epi <- readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/snp_epi.rds")
SNP_data = readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/dosage_8k.rds")
# change colnames of SNP_data 

```


## subset data to UKB data 

```{r}
snp_epi = snp_epi[snp_epi$study == "UKB", ]
SNP_data = SNP_data[SNP_data$study == "UKB", ]
```


# merge SNP data with epi data 

```{r}
sum(snp_epi$vcfid != SNP_data$vcfid)
# 0 -- ids match 

UKB_CRC_data = cbind(snp_epi[, 1:19], SNP_data[, -c(1:2)])

# uncomment if want to save ---
#saveRDS(UKB_CRC_data, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_dataset.rds")
```


```{r}
UKB_CRC_data = readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_dataset.rds")
```


Separate into Epi data and SNP data (from UKB dataset)

```{r}
epi_data = UKB_CRC_data[1:19]
SNP_data = UKB_CRC_data[-c(1:19)]
```


# Begin Data Analysis 
# look at the exposure distirbutions -- check for missingness 

```{r}
apply(epi_data[, c("outc", "age_ref_imp", "study", "asp_ref", 
                   "smk_ever", "alcmod", "alcheavy", "male")], MARGIN = 2, FUN = table, useNA = "always")
```


SNP == 2 isnt always the most rare group

make a params file with the SNP numbers for each sim 

```{r}
SNP_range = (1:ncol(SNP_data)) %>% matrix(ncol = 1)

write.table(SNP_range, file = "/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/param_file.txt", 
            row.names = F, col.names = F)

```

make 4 params files
```{r}
SNP_range = (1:ncol(SNP_data)) %>% matrix(ncol = 1)
for (i in 1:4)
{
  start_idx = 2000*(i-1)+1
  end_idx = 2000*(i)
  if(i == 4)
  {
    end_idx = length(SNP_range)
  }
 write.table(SNP_range[start_idx:end_idx, ,drop = F], file = paste0("/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/param_file_",i,".txt"), 
            row.names = F, col.names = F) 
}

```


Create the smoking dataset 

```{r}
UKB_CRC_smk = cbind(epi_data[c(1:7, 9, 16,19)], SNP_data) %>% na.omit
dim(UKB_CRC_smk)
saveRDS(UKB_CRC_smk, "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_smk.rds")
```

## Table for confounder distribution

```{r}
library(Hmisc)
library(table1)
ukb_table_1 = readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_smk.rds")
ukb_table_1$outc = factor(ukb_table_1$outc, levels = c(0, 1), labels = c("Control", "Case"))
ukb_table_1$male = factor(ukb_table_1$male, levels = c(0, 1), labels = c("Female", "Male"))
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}



label(ukb_table_1$male) <- "Male"
label(ukb_table_1$smk_ever) <- "Ever Smoker"
label(ukb_table_1$age_ref_imp) <- "Age at reference imputation (years)"
label(ukb_table_1$bmi10) <- "BMI (kg/m²)"
label(ukb_table_1$outc) <- "CRC Status"

my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=3), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}

table1(~ male + smk_ever + age_ref_imp + bmi10 | outc, 
       data = ukb_table_1, extra.col=list(`P-value`=pvalue), overall = F, 
       render.continuous=my.render.cont)


```

## Filter for SNPs -- apply after analysis 

```{r}
#UKB_CRC_smk = readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_smk.rds")
epi_data = UKB_CRC_smk[1:10]
SNP_data = UKB_CRC_smk[, -c(1:10)]
# do not exclude SNPs that have only levels 0 1 
# change cell count to 10 
# run SNPs as confounders with dosage adjustment (continuous)
rm_snps = apply(SNP_data, MARGIN = 2, FUN = function(x)
{
  tab = table(epi_data$smk_ever, epi_data$outc, round(x))
  
  # this is to safeguard against SNPs that do not have level 2 
  if(dim(tab)[3] < 3)
  {
    return(TRUE)
  }else
  {
  for(i in 1:dim(tab)[3])
  {
    if(sum(tab[,,i] < 20) > 0)
    {
      return(TRUE)
    }
  }
  return(FALSE)
  }
    
  }) %>% which
length(rm_snps)
```


## Global propensity scores for smoking 

We compute the global propensity scores independently so that they do not have to be independently estimated on each iteration, since run on different computing units.  
```{r}
UKB_CRC_smk = readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_smk.rds")

exposure = "smk_ever"
confounders = c("age_ref_imp", "pc1", "pc2","pc3", "bmi10", "male")

# find obs.weights 
outcome_table = table(UKB_CRC_smk$outc)
J = outcome_table[1] / outcome_table[2]
disease_prevalence = 2523 / (2523 + 481588)

# weight will be q0 for Y = 1 (cases) and (1-q0)*(1/J) for Y = 0 (controls)
# where q0 is the disease prevalence
obs.weights = UKB_CRC_smk$outc*disease_prevalence  + (1-UKB_CRC_smk$outc)*(1 - disease_prevalence)*(1/J)


smk_propensity_scores <- generate_propensity_SL(exposure_data = UKB_CRC_smk[c(exposure, confounders)], 
                                            obs.weights = obs.weights, 
                                            SL.library = c("SL.glmnet", "SL.rpart", "SL.bartMachine"), 
                                            SL.cvControl = list(V = 10L, stratifyCV = T, shuffle = TRUE, 
                                                                validRows = NULL), 
                                            parallel = T, 
                                            ncores = 10) %>% as.numeric

#saveRDS(smk_propensity_scores, file = "~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/smk_propensities.rds")

```

## tlGxE run on cluster -- output results here.  

```{r}
dir_path = "/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_smk_results_glmnet_no_hier_1-28-25"
files = list.files(dir_path)
```

## Process all SNPs

```{r}
SNP_files = list.files(dir_path)
result_list = list()

SNP_names = gsub(".RDS", "", gsub("SNP_", "", SNP_files))

for(i in 1:length(SNP_files))
{
  result_list[[i]] = readRDS(paste0(dir_path, "/",SNP_files[i]))
}

result_list = do.call(cbind, result_list)
colnames(result_list) = SNP_names

# Filter from cross tabulation criteria earlier 
result_list = result_list[, !(SNP_names %in% names(rm_snps))]

# Filter from NAs
if(length(which(colSums(is.na(result_list)) > 0)) > 0)
{
 result_list = result_list[, -which(colSums(is.na(result_list)) > 0)] 
}


level = 5*10^-4

list(
  ATE_aov = result_list[, which(result_list["ATE_EM_pvalue", ] < level), drop = F] %>% round(8), 
  ATE_1df = result_list[, which(result_list["ATE_EM_lin_pvalue", ] <level), drop = F] %>% round(8), 
  MOR_aov = result_list[, which(result_list["MOR_EM_pvalue", ] < level), drop = F] %>% round(8), 
  MOR_1df = result_list[, which(result_list["MOR_EM_mult_pvalue", ] <level), drop = F] %>% round(8)
)
```

## Make Latex table of results -- corresponds to Table 1 (upper)
```{r}

results_for_table = result_list[, which(result_list["ATE_EM_pvalue", ] < level |
                                          result_list["ATE_EM_lin_pvalue", ] <level |
                                          result_list["MOR_EM_pvalue", ] < level | 
                                          result_list["MOR_EM_mult_pvalue", ] <level
                                          ), drop = F] %>% t
# add the SNP the RERI found 
results_for_table = rbind(results_for_table, "X14.54415291" = result_list[, "X14.54415291"])
# reorder for printing
results_for_table = results_for_table[order(rownames(results_for_table)), ]

snp_bp_chr = rownames(results_for_table)
results_for_table = results_for_table[, - which(colnames(results_for_table) %in% c("var_ATE_E0", "var_ATE_E1",
                                                                                   "var_ATE_E2", "var_MOR_E0", 
                                                                                   "var_MOR_E1","var_MOR_E2", 
                                                                                   "ATE_EM_F_statistic",
                                                                                   "MOR_EM_F_statistic", 
                                                                                   "ATE_EM_lin_baseline_est", 
                                                                                   "ATE_EM_lin_est", 
                                                                                   "MOR_EM_mult_baseline_est", 
                                                                                   "MOR_EM_mult_est"))] %>% data.frame()

rs_ids = c("rs4360494", "rs61776719", "rs1003563", "rs11157993", "rs11623717","rs12898159", "rs2379128")
rownames(results_for_table) = rs_ids
colnames(results_for_table) = c("ATE$_{00}$", "ATE$_{01}$", "ATE$_{11}$",
                                "MOR$_{00}$", "MOR$_{01}$", "MOR$_{11}$", 
                                "ATE Codominant p", "MOR codominant p", 
                                "ATE additive p", "MOR additive p")

results_for_table = results_for_table %>% round(4)
chr = as.integer(sub("\\..*", "", gsub("X", "", snp_bp_chr)))
bp = as.integer(sub(".*\\.", "",  gsub("X", "", snp_bp_chr))) 

results_for_table = cbind(rs_ids,chr, bp, results_for_table)

latex_table = ""
p_thresh = 5*10^-4
for(i in 1:nrow(results_for_table))
{
  latex_table = paste(latex_table, "& & & ", paste(results_for_table[i, c(4:6)] %>% round(4), collapse = " & " ))
  for(j in c(10,12))
  {
    if(as.numeric(results_for_table[i, j]) <= p_thresh)
    {
      latex_table = paste(latex_table, "& \\textbf{", results_for_table[i, j] %>% round(4), "}") 
      
    }else
    {
      latex_table = paste(latex_table, "&" , results_for_table[i, j] %>% round(4)) 
    }
  }
  
  latex_table = paste(latex_table, "\\\\ \\textbf{",results_for_table[i,1], "}")
  latex_table = paste(latex_table, "&", paste(results_for_table[i, c(2, 3, 7:9)] %>% round(2), collapse = " & "))
   for(j in c(11,13))
  {
    if(as.numeric(results_for_table[i, j]) <= p_thresh)
    {
      latex_table = paste(latex_table, "& \\textbf{", results_for_table[i, j] %>% round(4), "}") 
      
    }else
    {
      latex_table = paste(latex_table, "&" , results_for_table[i, j] %>% round(4)) 
    }
  }
  latex_table = paste(latex_table, " \\\\ \\hline")
}

# can use this if want to change sci not
convert_scientific_to_latex <- function(text) {
  gsub("([0-9]+)e([-+]?)0*([1-9]\\d*)", "$\\1 \\\\times 10^{\\2\\3}$", text, perl = TRUE)
}

cat(latex_table)
```



## Function to perform manhattan plotting 

```{r}
library(qqman)
library(calibrate) # for textxy in self built manhattan function
library(latex2exp)

manhattan  = function (x, chr = "CHR", bp = "BP", p = "P", snp = "SNP", col = c("gray10", 
    "gray60"), chrlabs = NULL, suggestiveline = -log10(0.00001), 
    genomewideline = -log10(0.00000005), highlight = NULL, logp = TRUE, 
    annotatePval = NULL, annotateTop = TRUE,txt_size = 1, ...) 
{
  CHR = BP = P = index = NULL
  if (!(chr %in% names(x))) 
    stop(paste("Column", chr, "not found!"))
  if (!(bp %in% names(x))) 
    stop(paste("Column", bp, "not found!"))
  if (!(p %in% names(x))) 
    stop(paste("Column", p, "not found!"))
  if (!(snp %in% names(x))) 
    warning(paste("No SNP column found. OK unless you're trying to highlight."))
  if (!is.numeric(x[[chr]])) 
    stop(paste(chr, "column should be numeric. Do you have 'X', 'Y', 'MT', etc? If so change to numbers and try again."))
  if (!is.numeric(x[[bp]])) 
    stop(paste(bp, "column should be numeric."))
  if (!is.numeric(x[[p]])) 
    stop(paste(p, "column should be numeric."))
  if (!is.null(x[[snp]])) 
    d = data.frame(CHR = x[[chr]], BP = x[[bp]], P = x[[p]], 
                   pos = NA, index = NA, SNP = x[[snp]], stringsAsFactors = FALSE)
  else 
    d = data.frame(CHR = x[[chr]], BP = x[[bp]], P = x[[p]], 
                      pos = NA, index = NA)
  d <- d[order(d$CHR, d$BP), ]
  if (logp) {
    d$logp <- -log10(d$P)
  }
  else {
    d$logp <- d$P
  }
  d$index = rep.int(seq_along(unique(d$CHR)), times = tapply(d$SNP, 
                                                             d$CHR, length))
  nchr = length(unique(d$CHR))
  if (nchr == 1) {
    d$pos = d$BP
    xlabel = paste("Chromosome", unique(d$CHR), "position")
  }
  else {
    lastbase = 0
    ticks = NULL
    for (i in unique(d$index)) {
      if (i == 1) {
        d[d$index == i, ]$pos = d[d$index == i, ]$BP
      }
      else {
        lastbase = lastbase + max(d[d$index == (i - 1), 
                                    "BP"])
        d[d$index == i, "BP"] = d[d$index == i, "BP"] - 
          min(d[d$index == i, "BP"]) + 1
        d[d$index == i, "pos"] = d[d$index == i, "BP"] + 
          lastbase
      }
    }
    ticks <- tapply(d$pos, d$index, quantile, probs = 0.5)
    xlabel = "Chromosome"
    labs <- unique(d$CHR)
  }
  xmax = ceiling(max(d$pos) * 1.03)
  xmin = floor(max(d$pos) * -0.03)
  def_args <- list(xaxt = "n", bty = "n", xaxs = "i", yaxs = "i", 
                   las = 1, pch = 20, xlim = c(xmin, xmax), ylim = c(0, 
                                                                     ceiling(max(d$logp))), xlab = xlabel, ylab = expression(-log[10](italic(p))))
  dotargs <- list(...)
  do.call("plot", c(NA, dotargs, def_args[!names(def_args) %in% 
                                            names(dotargs)]))
  if (!is.null(chrlabs)) {
    if (is.character(chrlabs)) {
      if (length(chrlabs) == length(labs)) {
        labs <- chrlabs
      }
      else {
        warning("You're trying to specify chromosome labels but the number of labels != number of chromosomes.")
      }
    }
    else {
      warning("If you're trying to specify chromosome labels, chrlabs must be a character vector")
    }
  }
  if (nchr == 1) {
    axis(1, ...)
  }
  else {
    axis(1, at = ticks, labels = labs, ...)
  }
  col = rep_len(col, max(d$index))
  if (nchr == 1) {
    with(d, points(pos, logp, pch = 20, col = col[1], ...))
  }
  else {
    icol = 1
    for (i in unique(d$index)) {
      points(d[d$index == i, "pos"], d[d$index == i, "logp"], 
             col = col[icol], pch = 20, ...)
      icol = icol + 1
    }
  }
  if (suggestiveline) 
    abline(h = suggestiveline, col = "blue")
  if (genomewideline) 
    abline(h = genomewideline, col = "red")
  if (!is.null(highlight)) {
    if (any(!(highlight %in% d$SNP))) 
      warning("You're trying to highlight SNPs that don't exist in your results.")
    d.highlight = d[which(d$SNP %in% highlight), ]
    with(d.highlight, points(pos, logp, col = "green3", pch = 20, 
                             ...))
  }
  if (!is.null(annotatePval)) {
    if (logp) {
      topHits = subset(d, P <= annotatePval)
    }
    else topHits = subset(d, P >= annotatePval)
    par(xpd = TRUE)
    if (annotateTop == FALSE) {
      if (logp) {
        with(subset(d, P <= annotatePval), textxy(pos, 
                                                  -log10(P), offset = 0.625, labs = topHits$SNP, 
                                                  cex = txt_size), ...)
      }
      else with(subset(d, P >= annotatePval), textxy(pos, 
                                                     P, offset = 0.625, labs = topHits$SNP, cex = txt_size), 
                ...)
    }
    else {
      topHits <- topHits[order(topHits$P), ]
      topSNPs <- NULL
      for (i in unique(topHits$CHR)) {
        chrSNPs <- topHits[topHits$CHR == i, ]
        topSNPs <- rbind(topSNPs, chrSNPs[1, ])
      }
      if(length(topSNPs) > 0)
      {
        if (logp) {
          textxy(topSNPs$pos, 
                 -log10(topSNPs$P) + .5 + runif(nrow(topSNPs), -.5, .5), 
                 labs = topSNPs$SNP, cex = txt_size, font = 2,...)
          # textxy(topSNPs$pos-1, -log10(topSNPs$P)+.25, offset = .10, 
          #        labs = topSNPs$SNP, cex = txt_size, font = 2, ...)
        }
        else
        {
          textxy(topSNPs$pos, topSNPs$P, offset = 0.625, 
                 labs = topSNPs$SNP, cex = txt_size, ...)
        }
      }
    }
  }
  par(xpd = FALSE)
}

SNP_name = colnames(result_list)
rs_ids = c("rs4360494", "rs61776719", "rs1003563", "rs11157993", "rs11623717", "rs2379128")
SNP_sig = c("X1.38455891",  "X1.38461319",  "X12.6424577",  "X14.54411696", "X14.54414132", "X20.60898505")
for(i in 1:length(SNP_sig))
{
  snp = SNP_sig[i]
  SNP_name[which(SNP_name == snp)] = rs_ids[i]
}

manhattab_plot_data = data.frame(SNP_name = SNP_name,
                                 chr = as.integer(sub("\\..*", "", gsub("X", "", colnames(result_list)))),
                                 bp = as.integer(sub(".*\\.", "",  gsub("X", "", colnames(result_list)))), 
                                 p_ate_aov = result_list["ATE_EM_pvalue", ], 
                                 p_ate_lin = result_list["ATE_EM_lin_pvalue", ], 
                                 p_mor_aov = result_list["MOR_EM_pvalue", ], 
                                 p_mor_mult = result_list["MOR_EM_mult_pvalue", ]
)
```

## Manhattan Plots for tlGxE 
```{r}
thresh = 5*10^-4
pdf(file = "/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/ukb_crc_man_plot.pdf", width = 10, height = 9)
par(mfrow = c(2,2), mar = c(5, 5, 4, 2))

set.seed(1)
ate_aov_plot = manhattan(x =manhattab_plot_data, chr = "chr", 
                         bp = "bp", p = "p_ate_aov", 
                         snp = "SNP_name", 
                         annotatePval = thresh, 
                         suggestiveline = FALSE, 
                         genomewideline = -1*log(thresh,10), 
                         col = c("blue4", "orange3"), 
                         main = "tlGxE ATE Codominant",
                         txt_size = 1.2, 
                         ylim = c(0,5), 
                         cex.main = 1.5,                 # Increase plot title size
                         cex.lab = 1.5,                  # Increase axis labels size
                         cex.axis = 1.2,                 # Increase axis numbers size
                         cex.sub = 1.2, 
                         offset = -.2
)
set.seed(8)
ate_lin_plot = manhattan(x =manhattab_plot_data, chr = "chr", 
                         bp = "bp", p = "p_ate_lin", 
                         snp = "SNP_name", 
                         annotatePval = thresh, 
                         suggestiveline = FALSE, 
                         genomewideline = -1*log(thresh,10), 
                         col = c("blue4", "orange3"),
                         main = "tlGxE ATE Additive", 
                         txt_size = 1.2, 
                         ylim = c(0,5), 
                         cex.main = 1.5,                 # Increase plot title size
                         cex.lab = 1.5,                  # Increase axis labels size
                         cex.axis = 1.2,                 # Increase axis numbers size
                         cex.sub = 1.2, 
                         offset = 0
)
set.seed(8)
mor_aov_plot = manhattan(x =manhattab_plot_data, chr = "chr", 
                       bp = "bp", p = "p_mor_aov", 
                         snp = "SNP_name", 
                         annotatePval = thresh, 
                         suggestiveline = FALSE, 
                         genomewideline = -1*log(thresh,10), 
                         col = c("blue4", "orange3"),
                         main = "tlGxE MOR Codominant", 
                         txt_size = 1.2, 
                         ylim = c(0,5), 
                         cex.main = 1.5,                 # Increase plot title size
                         cex.lab = 1.5,                  # Increase axis labels size
                         cex.axis = 1.2,                 # Increase axis numbers size
                         cex.sub = 1.2, 
                         offset = 0
)
mor_mult_plot = manhattan(x =manhattab_plot_data, chr = "chr", 
                          bp = "bp", p = "p_mor_mult", 
                          snp = "SNP_name", 
                          annotatePval = thresh, 
                          suggestiveline = FALSE, 
                          genomewideline = -1*log(thresh,10), 
                          col = c("blue4", "orange3"),
                          main = "tlGxE MOR Additive", 
                          txt_size = 1.2, 
                          ylim = c(0,5), 
                         cex.main = 1.5,                 # Increase plot title size
                         cex.lab = 1.5,                  # Increase axis labels size
                         cex.axis = 1.2,                 # Increase axis numbers size
                         cex.sub = 1.2   
)
dev.off()
```

## Code to perform RERI testing -- includes commented implementation of linear odds model 
```{r}
library(car)
# n = 10000
# G = rbinom(n = n, size = 2, prob = .3)
# E = rbinom(n = n, size = 1, prob = .5)
# C = cbind(rnorm(n), rnorm(n))
# 
# odds_pi_i = exp(.2)*(1 + 0*G*E + .5*E + .2*G + .2*C[,1] + .1*C[,2])
# pi_i = odds_pi_i / (1+odds_pi_i)
# Y = rbinom(length(pi_i), size = 1, pi_i)

# linear_odds_neg_ll = function(B,design_matrix, y_vec) # log likelihood
# {
#   odds_pi_i <- pmax(exp(B[1]) * (1 + design_matrix %*% matrix(B[-1], ncol = 1)), 1e-10)
#   pi_i = odds_pi_i / (1+odds_pi_i)
#   ll = ifelse(y_vec == 1, pi_i, 1-pi_i) %>% log %>% sum(na.rm = F)
#   return(-1*ll)
# } 

# linear_odds_neg_ll_reparam = function(B,design_matrix, y_vec) # log likelihood
# {
#   odds_pi_i = exp(B[1]  + log(pmax.int(1 + design_matrix %*% matrix(B[-1], ncol = 1), 1e-6)))
#   pi_i = odds_pi_i / (1+odds_pi_i)
#   ll = sum(log(pi_i[y_vec == 1])) + sum(log(1 - pi_i[y_vec == 0]))
#   return(-1*ll)
# } 
# 
# reri_model = function(Y, E, G, C)
# {
  # n = 14000
  # G = rbinom(n = n, size = 2, prob = .3)
  # E = rbinom(n = n, size = 1, prob = .5)
  # C = cbind(rnorm(n), rnorm(n))
  # 
  # odds_pi_i = exp(.2)*(1 + 0*G*E + .5*E + .2*G + .2*C[,1] + .1*C[,2])
  # pi_i = odds_pi_i / (1+odds_pi_i)
  # Y = rbinom(length(pi_i), size = 1, pi_i)
#   # start with additive test ##############################
#   design_matrix = as.matrix(cbind(E*G, E, G, C)) 
#   #         B0 BEG BE BG
#   B_init = c(0, 0, 0, 0,  rep(0,ncol(C)))
#   
#   add_RERI <- optim(B_init,linear_odds_neg_ll_reparam, 
#                    design_matrix = design_matrix, 
#                    method = "Nelder-Mead",
#                    y_vec = Y,
#                    hessian = T, 
#                    control = list(maxit = 5000))
#   
#   coef_vec = add_RERI$par
#   names(coef_vec) = paste0("b", 1:length(coef_vec))
#   
#   RERI_add_effect = add_RERI$par[2]
#   RERI_add_cond_exp_effect = c(add_RERI$par[3], add_RERI$par[3] + RERI_add_effect, add_RERI$par[3] + 2*RERI_add_effect )
#   RERI_add_vcov = Matrix::nearPD(MASS::ginv(add_RERI$hessian))$mat
#   rownames(RERI_add_vcov) = names(coef_vec); colnames(RERI_add_vcov) = names(coef_vec)
#   RERI_add_var = RERI_add_vcov[2,2]
#   RERI_Z = RERI_add_effect / sqrt(RERI_add_var)
#   RERI_add_pvalue = 2*(1-pnorm(abs(RERI_Z)))
#   RERI_add_se = c(
#     deltaMethod(coef_vec, "b3", vcov = RERI_add_vcov)$SE, 
#     deltaMethod(coef_vec, "b3 + b2", vcov = RERI_add_vcov)$SE, 
#     deltaMethod(coef_vec, "b3 + 2*b2", vcov = RERI_add_vcov)$SE
#   )
#   ############################################################
#   # Codominant RERI Test 
#   
#   # full model 
#   G = round(G)
#   design_matrix = as.matrix(cbind(E*(G == 1), E*(G == 2), E, 1*(G == 1), 1*(G == 2),  C)) 
#   B_init = rep(0, ncol(design_matrix) + 1)
#   
#   CD_full_RERI <- optim(B_init,linear_odds_neg_ll_reparam, 
#                    design_matrix = design_matrix, 
#                    method = "Nelder-Mead",
#                    y_vec = Y,
#                    hessian = T, 
#                    control = list(maxit = 5000))
#   
#   ll_CD_full = -1*CD_full_RERI$value
#   
#   # nested model 
#   design_matrix = as.matrix(cbind(E, 1*(G == 1), 1*(G == 2),  C)) 
#   B_init = rep(0, ncol(design_matrix) + 1)
#   CD_nested_RERI <- optim(B_init,linear_odds_neg_ll_reparam, 
#                    design_matrix = design_matrix, 
#                    method = "Nelder-Mead",
#                    y_vec = Y,
#                    hessian = T, 
#                    control = list(maxit = 5000))
#   
#   ll_CD_nested = -1*CD_nested_RERI$value
#   
#   chi_sq = -2*(ll_CD_nested - ll_CD_full)
#   
#   CD_chi_sq_p = 1 - pchisq(chi_sq, df = 2)
#   CD_chi_sq_p
#   
#   RERI_codominant_cond_exp_effect = c(CD_full_RERI$par[4], CD_full_RERI$par[4] + CD_full_RERI$par[2], 
#                                        CD_full_RERI$par[4] + CD_full_RERI$par[3])
#   RERI_codom_vcov = Matrix::nearPD(MASS::ginv(CD_full_RERI$hessian))$mat
#   coef_vec = CD_full_RERI$par
#   names(coef_vec) = paste0("b", 1:length(coef_vec))
#   colnames(RERI_codom_vcov) =  names(coef_vec); rownames(RERI_codom_vcov) =  names(coef_vec)
#   
#   RERI_codom_se = c(
#     deltaMethod(coef_vec, "b4", vcov = RERI_codom_vcov)$SE, 
#     deltaMethod(coef_vec, "b4 + b2", vcov = RERI_codom_vcov)$SE, 
#     deltaMethod(coef_vec, "b4 + b3", vcov = RERI_codom_vcov)$SE
#   )
#   
#   
#   
#   return(
#     list(RERI_add = c(RERI_add_effect = RERI_add_effect, RERI_add_var = RERI_add_var, 
#                       RERI_add_pvalue = RERI_add_pvalue, RERI_add_cond_exp_effect = RERI_add_cond_exp_effect, 
#                       RERI_add_se = RERI_add_se), 
#          RERI_cod = c(RERI_cod_p = CD_chi_sq_p, RERI_codominant_cond_effect = RERI_codominant_cond_exp_effect, 
#                       RERI_codom_se = RERI_codom_se))
#   )
# }

reri_model = function(Y, E, G, C)
{
  # n = 14000
  # G = rbinom(n = n, size = 2, prob = .3)
  # E = rbinom(n = n, size = 1, prob = .5)
  # C = cbind(rnorm(n), rnorm(n))
  # 
  # odds_pi_i = exp(.2)*(1 + 0*G*E + .5*E + .2*G + .2*C[,1] + .1*C[,2])
  # pi_i = odds_pi_i / (1+odds_pi_i)
  # Y = rbinom(length(pi_i), size = 1, pi_i)
  
  data = data.frame(Y = Y, E = E, G = G, C = C)
  colnames(data)[1:3] = c("Y", "E", "G")
  formula = Y ~ . + E:G
  
  glm_mult = glm(formula, data = data, family = "binomial")
  
  coef_vec = coef(glm_mult)
  cov_mat = vcov(glm_mult)
  
  # G = 1
  P10 = "exp(G)"
  P01 = "exp(E)"
  P11 = "exp(G + E + `E:G`)"
  
  RERI_1 = paste0(P11,"  - ",P10," - ",P01," + 1")
  
  # G = 2
  P10 = "exp(2*G)"
  P01 = "exp(E)"
  P11 = "exp(2*G + E + 2*`E:G`)"
  
  RERI_2 =  paste0(P11,"  - ",P10," - ",P01," + 1")
  
  
  
  RERI_effects = list(
    RERI_G_1 = deltaMethod(coef_vec, RERI_1, vcov = cov_mat), 
    RERI_G_2 = deltaMethod(coef_vec, RERI_2, vcov = cov_mat)
  )
  
  RERI_add_effect = RERI_effects$RERI_G_1$Estimate
  RERI_add_cond_exp_effect = c(RERI_effects$RERI_G_1$Estimate, RERI_effects$RERI_G_2$Estimate)
  RERI_Z = RERI_add_effect /  RERI_effects$RERI_G_1$SE
  RERI_add_pvalue = 2*(1-pnorm(abs(RERI_Z)))
  RERI_add_se = c(
    RERI_effects$RERI_G_1$SE, RERI_effects$RERI_G_2$SE
  )
  
  return(
    list(RERI_add = c(RERI_add_effect = RERI_add_effect, 
                      RERI_add_pvalue = RERI_add_pvalue, RERI_add_cond_exp_effect = RERI_add_cond_exp_effect, 
                      RERI_add_se = RERI_add_se))
  )
}


```

# Perform Interaction Scan for Smoking using Standard GxE and RERI 

```{r}
library(car) # for delta method 
UKB_CRC_smk =readRDS("~/Desktop/USC/Targeted Learning/Simulations_Manuscript/UKB_analysis/OneDrive_1_10-30-2024/UKB_CRC_smk.rds")

smk_results = list()
SNP_names = setdiff(colnames(UKB_CRC_smk)[-c(1:10)], names(rm_snps))
outcome = "outc"
exposure = "smk_ever"
confounders = c("age_ref_imp", "pc1", "pc2","pc3", "bmi10", "male")
confounder_formula = paste(outcome, "~", paste(confounders, collapse = "+"))

C = model.matrix(formula(paste("~", paste(confounders, collapse = "+"))), data = UKB_CRC_smk)[,-1]

options(future.globals.maxSize = 8000 * 1024^2)
plan(multisession, workers = 10)

# GWIS performed here! 
smk_results = future_lapply(1:length(SNP_names), FUN = function(i)
{
  curr_SNP = SNP_names[i]
  interaction_term = paste(exposure, "*", curr_SNP)
  model_formula = paste(confounder_formula, "+", interaction_term) %>% formula
  
  # fit glm additive 
  gwis_model = glm(formula = model_formula, family = "binomial", data = UKB_CRC_smk)
  
  # 1df multiplicative 
  mult_int_term = paste0(exposure, ":", curr_SNP)
  mult_gxe_effect = gwis_model$coefficients[mult_int_term] 
  mult_gxe_pvalue = summary(gwis_model)$coefficients[mult_int_term, 4]
  mult_gxe_additive_COR = exp(c(coef(gwis_model)[exposure],coef(gwis_model)[exposure] + mult_gxe_effect, coef(gwis_model)[exposure] + 2*mult_gxe_effect))
  cov_mat = vcov(gwis_model)
  mult_gxe_additive_COR_se = c(deltaMethod(coef(gwis_model), "exp(smk_ever)", cov_mat)$SE, 
                               deltaMethod(coef(gwis_model), paste0("exp(smk_ever + `",exposure, ":",curr_SNP, "`)"), cov_mat)$SE, 
                                deltaMethod(coef(gwis_model), paste0("exp(smk_ever + 2*`",exposure, ":",curr_SNP, "`)"), cov_mat)$SE           
                               )
  
  # fit glm codominant 
  curr_SNP = SNP_names[i]
  model_formula_full = paste(confounder_formula,"+", exposure, "*I(round(", curr_SNP, ") == 1) + ",exposure, "*I(round(", curr_SNP, ") == 2) + ",
                        exposure, "+", "I(round(", curr_SNP, ") == 1) + ", "I(round(", curr_SNP, ") == 2)") %>% formula
  
  model_formula_nested = paste(confounder_formula,"+", exposure, "+", "I(round(", curr_SNP, ") == 1) + I(round(", curr_SNP, ") == 2)") %>% formula
  
  full_mod = glm(formula = model_formula_full, family = "binomial", data = UKB_CRC_smk)
  nested_mod = glm(formula = model_formula_nested, family = "binomial", data = UKB_CRC_smk)
  
  mult_codom_chisq = -2*(logLik(nested_mod) - logLik(full_mod))
  mult_codom_p = 1 - pchisq(mult_codom_chisq, df = 2)
  mult_codom_COR = exp(c(coef(full_mod)[exposure],
                         coef(full_mod)[exposure] + coef(full_mod)[paste0("", exposure, ":I(round(", curr_SNP, ") == 1)TRUE")], 
                         coef(full_mod)[exposure] + coef(full_mod)[paste0("", exposure, ":I(round(", curr_SNP, ") == 2)TRUE")]))
  cov_mat = vcov(full_mod)
  mult_codom_COR_se = c(deltaMethod(coef(full_mod), "exp(smk_ever)", cov_mat)$SE, 
                        deltaMethod(coef(full_mod), paste0("exp(smk_ever + `", exposure, ":I(round(", curr_SNP, ") == 1)TRUE`)"), cov_mat)$SE, 
                        deltaMethod(coef(full_mod), paste0("exp(smk_ever + `", exposure, ":I(round(", curr_SNP, ") == 2)TRUE`)"), cov_mat)$SE
                        )
  # RERI Testing 
  
  reri_testing = reri_model(Y = UKB_CRC_smk$outc, E = UKB_CRC_smk$smk_ever, 
                            G = UKB_CRC_smk[curr_SNP], C = C)
  
  
  gwis_results = c(mult_add_effect = mult_gxe_effect, 
                   mult_add_pvalue = mult_gxe_pvalue, 
                   mult_add_CORs = mult_gxe_additive_COR, 
                   mult_add_COR_se = mult_gxe_additive_COR_se, 
                   mult_codom_pvalue = mult_codom_p, 
                   mult_codom_CORs = mult_codom_COR, 
                   mult_codom_COR_se = mult_codom_COR_se, 
                   reri_add_effect = reri_testing$RERI_add["RERI_add_effect"],
                   reri_add_pvalue = reri_testing$RERI_add["RERI_add_pvalue"],
                   reri_add_CORs = reri_testing$RERI_add[3:4], 
                   reri_add_COR_se =  reri_testing$RERI_add[5:6])
  
  return(gwis_results)
  
})

plan(sequential)

# saveRDS(smk_results, "/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/standard_methods_scan.RDS")
# smk_results = readRDS("/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/standard_methods_scan.RDS")
smk_SNP_results = do.call(rbind,smk_results)
colnames(smk_SNP_results) = sub("\\..*", "", colnames(smk_SNP_results))
rownames(smk_SNP_results) = SNP_names
smk_SNP_results = data.frame(smk_SNP_results)


sig_thresh = 5*10^-4
list(
  GxE_mult_add = smk_SNP_results[which(smk_SNP_results[,"mult_add_pvalue"] < sig_thresh), , drop = F] %>% round(8), 
  GxE_mult_codom = smk_SNP_results[which(smk_SNP_results[,"mult_codom_pvalue" ] <sig_thresh), , drop = F] %>% round(8), 
  RERI_add = smk_SNP_results[which(smk_SNP_results[,"reri_add_pvalue" ] < sig_thresh), , drop = F] %>% round(8)
)

smk_SNP_results$chr = as.integer(sub("\\..*", "", gsub("X", "", rownames(smk_SNP_results))))
smk_SNP_results$bp = as.integer(sub(".*\\.", "",  gsub("X", "", rownames(smk_SNP_results))))
smk_SNP_results$SNP_names = rownames(smk_SNP_results)


smk_SNP_results$SNP_names[which(smk_SNP_results$SNP_names == "X1.38461319")] = "rs61776719"
smk_SNP_results$SNP_names[which(smk_SNP_results$SNP_names == "X12.6424577")] = "rs1003563"
smk_SNP_results$SNP_names[which(smk_SNP_results$SNP_names == "X14.54414132")] = "rs11623717"
smk_SNP_results$SNP_names[which(smk_SNP_results$SNP_names == "X20.60898505")] = "rs2379128"

# rename significant SNPs with RSIDs 

```


# Latex Table for Standard Methods -- Table 2

```{r}
results_for_table = smk_SNP_results[which(smk_SNP_results[, "mult_add_pvalue" ] < level |
                                          smk_SNP_results[, "mult_codom_pvalue"] <level |
                                          smk_SNP_results[, "reri_add_pvalue"] < level), , drop = F] 
snp_bp_chr = rownames(results_for_table)

rs_ids = c("rs4360494", "rs61776719", "rs1003563", "rs11157993", "rs11623717", "rs12898159", "rs2379128")
rownames(results_for_table) = rs_ids
results_for_table = results_for_table[-24]

chr = as.integer(sub("\\..*", "", gsub("X", "", snp_bp_chr)))
bp = as.integer(sub(".*\\.", "",  gsub("X", "", snp_bp_chr))) 

results_for_table = cbind(rs_ids,chr, bp, results_for_table)

latex_table = ""
p_thresh = 5*10^-4
for(i in 1:nrow(results_for_table))
{
  latex_table = paste(latex_table, "& & & &", paste(results_for_table[i, c(21:22)] %>% round(3), collapse = " & " ))
  for(j in c(20))
  {
    if(as.numeric(results_for_table[i, j]) <= p_thresh)
    {
      latex_table = paste(latex_table, "& & \\textbf{", results_for_table[i, j] %>% round(4), "}") 
      
    }else
    {
      latex_table = paste(latex_table, "& &" , results_for_table[i, j] %>% round(4)) 
    }
  }
  
  latex_table = paste(latex_table, "\\\\ \\textbf{",results_for_table[i,1], "}")
  latex_table = paste(latex_table, "&", paste(results_for_table[i, c(2, 3, 13:15)] %>% round(2), collapse = " & "))
   for(j in c(12,5))
  {
    if(as.numeric(results_for_table[i, j]) <= p_thresh)
    {
      latex_table = paste(latex_table, "& \\textbf{", results_for_table[i, j] %>% round(4), "}") 
      
    }else
    {
      latex_table = paste(latex_table, "&" , results_for_table[i, j] %>% round(4)) 
    }
  }
  latex_table = paste(latex_table, " \\\\ \\hline")
}

cat(latex_table)
```


# RERI and Standard GxE Manhattan Plots

```{r}
# results for RERI and MULT GxE 
# this seed is for the random jitter in the manhattan function 
thresh = 5*10^-4
pdf(file = "/Users/danielrud/Desktop/USC/Targeted Learning/Simulations_Manuscript/ukb_crc_gxe_mult_reri.pdf", width = 10, height = 9)
layout(matrix(c(1,1, 2,2, 0, 3, 3, 0), nrow = 2, byrow = TRUE))
par( mar = c(5, 5, 4, 2))

set.seed(1)
gxe_codom_plot = manhattan(x =smk_SNP_results, chr = "chr", 
                         bp = "bp", p = "mult_codom_pvalue", 
                         snp = "SNP_names", 
                         annotatePval = thresh, 
                         suggestiveline = FALSE, 
                         genomewideline = -1*log(thresh,10), 
                         col = c("blue4", "orange3"),
                         main = "GxE Codominant", 
                         txt_size = 1.5, 
                         ylim = c(0,5), 
                         cex.main = 2,                 # Increase plot title size
                         cex.lab = 2,                  # Increase axis labels size
                         cex.axis = 1.5,                 # Increase axis numbers size
                         cex.sub = 1.5, 
                         offset = .4)

gxe_add_plot = manhattan(x =smk_SNP_results, chr = "chr", 
                         bp = "bp", p = "mult_add_pvalue", 
                         snp = "SNP_names", 
                         annotatePval = thresh, 
                         suggestiveline = FALSE, 
                         genomewideline = -1*log(thresh,10), 
                         col = c("blue4", "orange3"),
                         main = "GxE Additive", 
                         txt_size = 1.5, 
                         ylim = c(0,5), 
                         cex.main = 2,                 # Increase plot title size
                         cex.lab = 2,                  # Increase axis labels size
                         cex.axis = 1.5,                 # Increase axis numbers size
                         cex.sub = 1.5)

#22
set.seed(24)
reri_add_plot = manhattan(x =smk_SNP_results, chr = "chr", 
                         bp = "bp", p = "reri_add_pvalue", 
                         snp = "SNP_names", 
                         annotatePval = thresh, 
                         suggestiveline = FALSE, 
                         genomewideline = -1*log(thresh,10), 
                         col = c("blue4", "orange3"),
                         main = "RERI", 
                        txt_size = 1.5, 
                         ylim = c(0,5), 
                         cex.main = 2,                 # Increase plot title size
                         cex.lab = 2,                  # Increase axis labels size
                         cex.axis = 1.5,                 # Increase axis numbers size
                         cex.sub = 1.5, 
                         offset = -.1)


dev.off()
```















